---
title: "MA [46]15: Lab activity 1"
author: "Your Name"
output: html_document
---

# Introduction

In this lab activity we do a preliminary analysis of global social and economic indices. We use the *Social Progress Index* (SPI) from 2017 and the *Gross Domestic Product* (GDP) per capita from 2018. We downloaded the [social](https://www.socialprogressindex.com/assets/downloads/2017-Results.xlsx.zip) and [economic](http://api.worldbank.org/v2/en/indicator/NY.GDP.PCAP.PP.CD?downloadformat=excel) index data and put them at the `data` folder. We can now load them, along with the `tidyverse` package: 

```{r init, warning = FALSE}
library(tidyverse)
spi <- readxl::read_xlsx("2017 Results.xlsx") %>%
  mutate(spi = `Social Progress Index`)
ief <- readxl::read_xls("index2018_data.xls") %>%
  mutate(score = as.numeric(`2018 Score`), # Index of Economic Freedom (IEF)
         gdp = as.numeric(`GDP per Capita (PPP)`),
         Region = factor(Region))
```

We can now join the two datasets using `Country`:

```{r join, message = FALSE}
(ief.spi <- left_join(select(spi, Country, spi),
                     select(ief, Country, Region, score, gdp)))
```


# Visualization

Our first visualization is SPI against GDP:

```{r spi-gdp, warning = FALSE}
ggplot(ief.spi, aes(gdp, spi)) + geom_point() +
  labs(x = "2018 GDP per capita (PPP)", y = "2017 Social progress index")
```

Our next visualization plots SPI against IEF:

```{r spi-ief, warning = FALSE}
ggplot(ief.spi, aes(gdp, spi)) + geom_point() +
  labs(x = "2018 Index of economic freedom", y = "2017 Social progress index")
```

# Discussion